---
title: "Biostat M280 Homework 2"
subtitle: Due Feb 8 @ 11:59PM
author: "Edward Yu, UID: 705241661"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Master Build Status
[![Build Status](https://travis-ci.com/edwardmjyu/biostat-m280-2019-winter.svg?token=yQzFps26vPmcxdEGm582&branch=master)](https://travis-ci.com/edwardmjyu/biostat-m280-2019-winter)  

Develop Build Status
[![Build Status](https://travis-ci.com/edwardmjyu/biostat-m280-2019-winter.svg?token=yQzFps26vPmcxdEGm582&branch=develop)](https://travis-ci.com/edwardmjyu/biostat-m280-2019-winter)
  
## Q2. FizzBuzz Test

My FizzBuzz function has a two step verification process. 

1. The first pass checks whether the user input is a numeric scalar or vector; if it is not, then the function returns a printed message notifying the user of a Type Error and asks the user to input integers only. 

2. The second verification is a more refined, checking whether the input is `NA`, `Inf`, or a decimal value. This function was primarily created from the need to defensively code since a user can pass a variety of values into the function. The `FizzBuzz` function first checks itself if the input is numeric. This first step is necessary because if a character value is passed through the function, the character will coerce the rest of the vector into a character vector since R objects cannot be of multiple classes unless it is defined as a list. 

3. Next, the for loop goes through each element of the user input and checks whether any violations occur via calling the `invalid_arg` function. If the check passes, then the element enters the meat of the `FizzBuzz` function. The function must check whether it is divisible by both 5 and 3 first before other checks to make sure `FizzBuzz` is properly printed. 

```{r}
#Check arguments of input and determine whether input is valid
invalid_arg <- function(input) {
  if (is.numeric(input) == 0) {
    print("An input of the argument is a non-numeric, integers only.")
    return(0)
  } else if (is.na(input)) {
    print("An input of the argument is NA, integers only!")
    return(0)
  } else if (is.infinite(input)) {
    print("An input of the argument is infinity and beyond, integers only.")
    return(0)
  } else if (input %% 1 != 0) {
    print("An input of the argument is a non-integer, integers only.")
    return(0)
  } else {
    return(1)
  }
}

#FizzBuzz combines the first function testing args and outputs proper response.
FizzBuzz <- function(x) {
  if (!is.numeric(x)) {
    return ("Inputted scalar or vector has a Type Error, input integers only.")
  } 
  for (input in x) {
    check <- invalid_arg(input)
    if(check) {
      if (input %% 5 == 0 && input %% 3 == 0) {
        print("FizzBuzz")
      } else if (input %% 5 == 0) {
        print("Buzz")
      } else if (input %% 3 == 0) {
        print("Fizz") 
      } else {
        print(input)
      }
    }
  }
}
```

Here we test the function against several inputs:
```{r}
input1 <- 5 # scalar integer
input2 <- 1:15 # vector of integers
input3 <- "hello, M280 grader!" # character
input4 <- c(3, 5, 15, 3.14159265359) # mixture of numbers and characters
input5 <- -20 # negative values
input6 <- c(3, Inf) # Infinity value
input7 <- c(5, NA) # NA value

FizzBuzz(input1)
FizzBuzz(input2)
FizzBuzz(input3)
FizzBuzz(input4)
FizzBuzz(input5)
FizzBuzz(input6)
FizzBuzz(input7)
```


## Question 3. Rcpp

```{Rcpp}
#include <Rcpp.h>
#include <iostream>
using namespace Rcpp;
using namespace std;

// [[Rcpp::export]]

void Fizzy_cpp(NumericVector & nums){
  for(unsigned i = 0; i < nums.size(); ++i) {
    if (i % 5 == 0 && i % 3 == 0)
      std::cout << "FizzBuzz" << std::endl;
    else if (i % 3 == 0) 
      std::cout << "Fizz" << std::endl;
    else if (i % 5 == 0)
      std::cout << "Buzz" << std::endl;
    else 
       std::cout << i << std::endl;
  }
}
```


