---
title: "Biostat M280 Homework 3"
author: "Edward Yu"
subtitle: Due Mar 1 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use tidyverse and bash to explore following two data sets.

## Q1 LA City Employee Payroll

The `/home/m280data/la_payroll/City_Employee_Payroll.csv` file on teaching server contains payroll information of LA City employees in years 2013-2018. It was downloaded from [LA City Controller's Office](https://controllerdata.lacity.org/Payroll/City-Employee-Payroll/pazn-qyym). Make a Shiny app to facilitate visualization of this data. 

1. For efficiency of the Shiny app, you should first pre-process, pare down, tidy, and save the data, e.g., as a compressed RDS file, to be used in the app.

2. **Total payroll by LA City**. Visualize the total LA City payroll of each year, with breakdown into base pay, overtime pay, and other pay.

3. **Who earned most?** Visualize the payroll information (total payment with breakdown into base pay, overtime pay, and other pay, Department, Job Title) of the top $n$ highest paid LA City employees in a specific year. User specifies $n$ (default 10) and year (default 2017).

4. **Which departments earn most?** Visualize the mean or median payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ earning departments. User specifies $n$ (default 5), year (default 2017), and method (mean or median, default median).

5. **Which departments cost most?** Visualize the total payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ expensive departments. User specifies $n$ (default 5) and year (default 2017).

6. Visualize any other information you are interested in.

7. Publish your Shiny app to <https://www.shinyapps.io> and share the link.

**Solution**: Below is a link to my shiny new app:  <https://edwardmjyu.shinyapps.io/lapayroll_shiny/>

## Q2 LA City Parking War

The SQLite database `/home/m280data/la_parking/LA_Parking_Citations.sqlite` on teaching server contains information about parking tickets in LA City. It was downloaded from [LA Open Data Portal](https://data.lacity.org/A-Well-Run-City/Parking-Citations/wjz9-h9np). Connect to the database and answer following questions using plots and summary statistics. In this exercise, you are **not** allowed to load whole data into memory. Use the _transform in database, plot in R_ strategy.


```{r, results=FALSE}
### Initiate Libraries ###
library("DBI")
library("RSQLite")
library("tidyverse")
### Prepare SQL Connection ###
con <- dbConnect(RSQLite::SQLite(), 
                "/home/m280data/la_parking/LA_Parking_Citations.sqlite")
dbListTables(con) #Check tables available in connection
latix <- dplyr::tbl(con, "latix")
knitr::opts_chunk$set(connection = "con")
```

**1. How many tickets are in this data set? Which time period do these tickets span? Which years have most data?**
```{sql connection="con", results=FALSE}
SELECT count(DISTINCT Ticket_number) AS num_ticket FROM latix;
```
**Solution**: There are 7,656,362 distinct tickets, SQL removes any `NA` data.

```{sql connection="con"}
SELECT datetime(MIN(Issue_DateTime), 'unixepoch') AS "Start Time",
       datetime(MAX(Issue_DateTime), 'unixepoch') AS "End Time"
FROM latix;
```
**Solution**: The time period the tickets span is from `2015-07-02 01:00:00 UTC` to `2019-01-25 23:58:00 UTC`.

```{sql connection="con"}
SELECT 
  strftime('%Y', datetime(Issue_DateTime, 'unixepoch')) AS Year, 
  count(*) AS Count
FROM latix
GROUP BY 1
ORDER BY 2 DESC
```
**Solution**: The year with the most tickets was in 2017 with `2,235,243` tickets.

**2. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?**

Calculating for hour:
```{r}
hour <- 
  "SELECT 
    strftime('%H', DATETIME(Issue_DateTime, 'unixepoch')) AS Hour
  FROM latix
  WHERE Issue_DateTime IS NOT NULL"

hour <- dbGetQuery(con, hour) 
ggplot(hour, aes(Hour)) + geom_bar()
latix %>% 
  select(Issue_Hour) %>%
  filter(!is.na(Issue_Hour)) %>%
  group_by(Issue_Hour) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```
**Solution**: The bar chart above shows that `12pm` is the most likely time to receive a ticket while `5am` is the least likely time. 

Calculating for weekday:
```{r}
weekday <- 
  "SELECT 
    strftime('%w', DATETIME(Issue_DateTime, 'unixepoch')) AS Weekday
  FROM latix
  WHERE Issue_DateTime IS NOT NULL"

weekday <- dbGetQuery(con, weekday) 
ggplot(weekday, aes(Weekday)) + geom_bar()
latix %>% 
  select(Issue_Wday) %>%
  filter(!is.na(Issue_Wday)) %>%
  group_by(Issue_Wday) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```
**Solution**: The bar chart shows that most tickets are issued on `Monday`and least likely to get a ticket on `Saturday`. 

Calculating for day of month:
```{r}
dmonth <- 
  "SELECT 
    strftime('%d', DATETIME(Issue_DateTime, 'unixepoch')) AS Mday
  FROM latix
  WHERE Issue_DateTime IS NOT NULL"

mday <- dbGetQuery(con, dmonth) 
ggplot(mday, aes(Mday)) + geom_bar()
latix %>% 
  select(Issue_Day) %>%
  filter(!is.na(Issue_Day)) %>%
  group_by(Issue_Day) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```
**Solution**: The bar chart shows that it most likely to get a ticket on day `13` and least likely to get a ticket on day `31`. This is most likely due to there being the least number of months with `31` days. 

Calculating for month:
```{r}
month <- 
  "SELECT 
    strftime('%m', DATETIME(Issue_DateTime, 'unixepoch')) AS Month
  FROM latix
  WHERE Issue_DateTime IS NOT NULL"

month <- dbGetQuery(con, month) 
ggplot(month, aes(Month)) + geom_bar()
latix %>% 
  select(Issue_Month) %>%
  filter(!is.na(Issue_Month)) %>%
  group_by(Issue_Month) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```
**Solution**: The bar chart shows that it most likely to get a ticket in `August` and least likely to get a ticket in `February`. This is most likely due to there being the least days in `February`.

**3. Which car makes received most citations?**
```{r}
car <- 
 "SELECT
    Make AS Car,
    COUNT(*) AS Number
  FROM latix
  WHERE Make IS NOT NULL
  GROUP BY 1
  ORDER BY 2 DESC
  LIMIT 1"

dbGetQuery(con, car) 
```
**Solution**: The make that received the most citations was Toyota with a total of `1,270,804` citations.

**4. How many different colors of cars were ticketed? Which color attracted most tickets?**
```{r}
color <- 
 "SELECT
    Color AS Color,
    COUNT(*) AS Number
  FROM latix
  WHERE Color IS NOT NULL
  GROUP BY 1
  ORDER BY 2 DESC
  LIMIT 1"

color <- dbGetQuery(con, color) 
c(color[1,],nrow(color))
```
**Solution**: Black cars were flagged for the most number of tickets at `1,651,398` tickets.
There are a total of `103` different colors.

**5. What are the most common ticket types?**
```{r}
ticket_type <- 
 "SELECT
    Violation_Description AS violation,
    COUNT(*) AS Number
  FROM latix
  WHERE violation IS NOT NULL
  GROUP BY 1
  ORDER BY 2 DESC
  LIMIT 1"

dbGetQuery(con, ticket_type) 
```
**Solution**: The most type of ticket is `NO PARK/STREET CLEAN` with `2,234,699` tickets.

**6. How much money was collected on parking tickets in 2016, 2017 and 2018?**
```{r}
money <-   
  "SELECT
    strftime('%Y', DATETIME(Issue_DateTime, 'unixepoch')) AS Year,
    SUM(Fine_amount) AS money
  FROM latix
  WHERE Issue_DateTime IS NOT NULL
  GROUP BY 1"

money <- dbGetQuery(con, money)
money %>% 
  filter(Year %in% c("2016", "2017", "2018"))
```
**Solution**: 
The amount of money collected in tickets for the following years:
1. 2016: `$152,145,538`
2. 2017: `$157,122,489`	
3. 2018: `$138,875,787`

**7. If you've been ticketed in LA County, did you find your ticket in this data set?**

I've never been ticketed in LA before!